Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2019-12-01 xrb/paper.tex
More edits, discussed streamline plot
Still need to talk about rotation, and show Don's stuff. The initial
model and general features sections need a little bit more work.

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \rho \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the
pressure and $E$ is the specific total energy, which is related to the
specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
momentum equation includes gravity, described by gravitation
acceleration $\gb$ and rotational forces described by angular velocity
$\Omegab$, with $\rb$ the position vector from the origin.
Species are described by mass fractions, $X_k$ (such that $\sum_k X_k
= 1$), and creation rates, $\omegadot$, and are related to the total
specific energy generation rate, $\epsdot$\MarginPar{Is notation
  consistent between E, e, epsilondot, and enucdot?}.  The total mass
conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Thermal diffusion is 
described by a thermal conductivity $\kth$ and temperature $T$.

@@

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega_0 u
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:\MarginPar{In conjunction with total mass conservation?}
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

@@

\begin{figure}[t]
	\plotone{grids.pdf}
	\caption{\label{fig:grids}A slice through part of a 2D simulation showing the four-level
		grid structure. Note that the boxes shown are not the simulation zones, which span
		$10~\mathrm{cm}$ at the finest level, but subdomains containing approximately equal
		numbers of zones that are each managed by a separate MPI process. A coarse base grid
		extending to the upper boundary is drawn in white, with the green, blue, and black
		grids showing the jumps in refinement needed to fully resolve the fuel layer and
		underlying neutron star.}
\end{figure}

@@

{\color{red} We should decide on a notation for 2D and laminar flame speeds.}

The speed and thickness of a laminar helium flame are determined by the
energy generation rate and conductivity, and scale roughly as
\begin{equation}
\label{eq:flame_scaling}
v_f \approx \sqrt{\kth \dot{\epsilon}} \qquad
\lambda_f \approx \sqrt{\frac{\kth}{\dot{\epsilon}}}
\end{equation}
\citep{orourke:1979,khokhlov:1993}.
At the densities we consider in this simulation, the pure He flame
speed is quite slow and would require a prohibitively long integration
time to see significant evolution of the burning.  We therefore
consider boosted flames in this first paper, to accelerate the burning
and allow us to understand the qualitative effects of laterally
propagating flames.  To boost the flame while keeping the thickness
the same, we can multiply both the burning rate and the conductivity
by the same factor.  For our standard calculations, we choose 10 for
each, to give a 10$\times$ faster flame speed.  We call this the ``10/10''
flame.  We will also do a simulation with the reactions and
conductivity both boosted by 5, the ``5/5'' flame.

@@

We measure the flame width as:
\begin{equation}
\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}
Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

@@

To assess the sensitivity of the flame propagation to the various approximations
we made, we ran a suite of simulations. 
Table~\ref{table:sim_names} summarizes these simulations. \MarginPar{This table has different model names from the flame speed table}
The majority of them used a reaction rate boosting of $10$ and a
conductivity boosting of $10$, which should increase the flame speed
by a factor of $10$.  Most simulations used a resolution of
$10~\mbox{cm}$ and a domain width of a little more than one kilometer.  
We use an artifically high rotation rate of
$2000~\mbox{Hz}$, which gives a Rossby length of
\begin{equation}
L_R = \frac{\sqrt{g H_0}}{\Omega} \sim 3\times 10^4~\mathrm{cm}
\end{equation}
using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
quarter of the domain width.  The run at $20~\mbox{cm}$ resolution used
one fewer level of refinement.  The slower rotating case using a
slightly wider domain to accommodate the expected larger Rossby
length.  We note that the entire simulation framework for these
calculations is freely available in the \castro\ github
repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
the simulation names defined in the table to refer to specific runs
and we'll use the 10/10 run as the reference calculation.

\begin{deluxetable}{lcccccc}
\tablecaption{\label{table:sim_names} Simulation parameters.}
\tablehead{\colhead{name} &
           \colhead{reaction} &
           \colhead{conductivity} &
           \colhead{fine grid} &
           \colhead{rotation} &
           \colhead{domain size} &
           \colhead{network} \\
%
           \colhead{} &
           \colhead{boost} &
           \colhead{boost} &
           \colhead{resolution} &
           \colhead{rate} &
           \colhead{} &
           \colhead{}
}
\startdata
10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
10/10-slow & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
10/10-lores & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
\enddata
\end{deluxetable}

@@

Our default resolution puts $\sim 80$ zones vertically in the
``cool'' \MarginPar{double check} model atmosphere height.  To
understand the effects of resolution, we also performed a run with one
fewer refinement level, giving a $20~\mbox{cm}$ resolution overall (this is our
10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
20~ms.  The structure is largely the same as the 10/10 run, with
largely the same flame shape and position, and the same structure in
the energy generation rate.  Since we do not have a jump in refinement
right below the atmosphere, we don't see the temperature feature from
the initial model mapping there, but we do see some cooling in the
\isot{Ni}{56} region.  This is likely a resolution effect, again due
to the strong gradient in temperature at the base of the atmosphere.
The strong agreement in the low resolution run to the 10/10 run gives
us confidence that we are capturing the flame physics properly.

@@

We also considered the effect of the network size.
Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
with the standard 13-isotope ``aprox13'' network and the reduced\MarginPar{We use {\tt tt} instead of quotes for the networks in the numerical approach section.}
7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
is slightly more advanced and has a higher $\bar{A}$ than the iso7
case.  We see in the next section that these two networks give largely
the same flame speed.  The reduced network size saves a lot of memory,
which will be useful when we transition to 3-d simulations.

\begin{figure}[t]
\centering
\plotone{network_compare}
\caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
  aprox13 vs a run with the iso7 network.}
\end{figure}

@@

To measure the propagation rate of the burning front, we track the
location beyond the hottest part of the flame\MarginPar{what does this mean?} where
$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of its
maximum value. Using $0.1 \%$ of the peak value rather than the peak value
itself helps to mitigate fluctuations caused by localized fluid
motions, as the reduced value lies in a less turbulent part of the
burning region. We also average $\dot{e}_\mathrm{nuc}$ over the  \MarginPar{I think we should discuss the averaging first, before we discuss the peak value}
vertical coordinate to further dampen these fluctuations. After an
initial transient period typically spanning $\sim 3$ ms, the flame
settles into a state of steady propagation. The position data here are
well fitted by a linear function of time (as seen in Figure
\ref{fig:flame_speed}), and the resulting slope gives the velocity of
the flame front.

\begin{deluxetable}{lc}
	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured in 2D calculations.}
	\tablehead{\colhead{run} & \colhead{$s_\mathrm{front}$ (km s$^{-1}$)}} 
	\startdata
	10/10 boost & $9.18 \pm 0.03$ \\
	5/5 boost & $4.00 \pm 0.01$ \\
	iso7 network & $7.56 \pm 0.02$ \\
	lowres & $9.33 \pm 0.04$ \\
	1000 Hz & $18.6 \pm 0.16$ \\
	\enddata
	%
\end{deluxetable}

Table \ref{table:flame_speeds_multid} gives the flame speed measured in each
multidimensional simulation, as well as laminar speeds obtained from
corresponding 1D runs. {\color{red} Should we include the laminar speeds again for reference?}
The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts. 
The increase in flame speed is likely a product of the larger flame surface area and
hydrodynamical effects such as turbulence, wrinkling, and convective cycles, which
bring cooler fuel from ahead of the front into the hottest part of the burning region.

@@

As we see in Figure~\ref{fig:flame_speed}, all of the various
approximations behave as we would expect.  The 1000~Hz run goes 2
times faster than the 2000~Hz run, as expected by the reduced Rossby
radius \citep{cavecchi:2013}).  The 5/5 boost run is half the speed of
the 10/10 boost run, also as expected, consistent with the 1D laminar
flames and behaving as predicted by Eq.~\ref{eq:flame_scaling}.  This
confirms that the role of the Coriolis force is really to boost the
laminar flame by the geometrical/hydrodynamical effect expected ($v_f
= v_l * L_R / H$), similarly to what the 1000 Hz simulation
said.

@@

\AssignTo{Alice, Don}

{\color{red} Show whatever it was that Don cooked up. Also need to discuss the streamline plot.}

Baroclinicity is calculated as
\begin{equation}
    \boldsymbol{\psi} = \frac{1}{\rho^2} \nablab p \times \nablab \rho,
\end{equation}
and shows the misalignment of the local density and pressure gradients
\citep{Malone2014a}. As we are considering a 2d \MarginPar{We've used
  at least three different ways of writing 2-d/2D/2d so far. These
  should be changed so we used one style throughout.}system, the
component of the baroclinicity we consider here (out of the plane)
reflects the misalignment of the fields in the plane of the
simulation. Figure~\ref{fig:baroclinicity} shows that the
baroclinicity peaks along the flame front. This baroclinicity
generates vorticity, which in turn entrains material along the surface
of the flame (see also \citet{cavecchi:2013}). In 3D, his same vorticity
should perturb the flame front and further increase the flame speed.
\citep{Cavecchi2019}.

@@

@@

\AssignTo{everyone}

@@

We use a 2-d axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
and Neumann boundaries on all other sides.  In this geometry, the solution behaves like
a spherical problem, and has the form:
\begin{equation}
  T(r,t) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
\end{equation}
where $t_0$ is a small time used to set the initial width of the
Gaussian, $r$ is the distance from the origin, and $D$ is the
diffusion coefficient.  $T_1$ is the ambient temperature and $T_2$ is
the temperature of the peak of the Gaussian.  We use $D = 1$, $T_1 = 1$, $T_2 = 2$,
and $t_0 = 0.001$.

>>>>

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \rho \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the
pressure and $E$ is the specific total energy, which is related to the
specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
momentum equation includes gravity, described by gravitation
acceleration $\gb$ and rotational forces described by angular velocity
$\Omegab$, with $\rb$ the position vector from the origin.
Species are described by mass fractions, $X_k$ (such that $\sum_k X_k
= 1$), and creation rates, $\omegadot$, and are related to the total
specific energy generation rate, $\epsdot$.  The total mass
conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Thermal diffusion is 
described by a thermal conductivity $\kth$ and temperature $T$.

@@

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega_0 u
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

@@

\begin{figure}[t]
	\plotone{grids.pdf}
	\caption{\label{fig:grids}A section of a 2D simulation showing the four-level
		grid structure. Note that the boxes shown are not the simulation zones, which span
		$10~\mathrm{cm}$ at the finest level, but subdomains containing approximately equal
		numbers of zones that are each managed by a separate MPI process. A coarse base grid
		extending to the upper boundary is drawn in white, with the green, blue, and black
		grids showing the jumps in refinement needed to fully resolve the fuel layer and
		underlying neutron star.}
\end{figure}

@@

The speed and thickness of a laminar helium flame are determined by the
energy generation rate and conductivity, and scale roughly as
\begin{equation}
\label{eq:flame_scaling}
s_L \approx \sqrt{\kth \epsdot} \qquad
\lambda_L \approx \sqrt{\frac{\kth}{\epsdot}}
\end{equation}
\citep{orourke:1979,khokhlov:1993}.
At the densities we consider in this simulation, the pure He flame
speed is quite slow and would require a prohibitively long integration
time to see significant evolution of the burning.  We therefore
consider boosted flames in this first paper, to accelerate the burning
and allow us to understand the qualitative effects of laterally
propagating flames.  To boost the flame while keeping the thickness
the same, we can multiply both the burning rate and the conductivity
by the same factor.  For our standard calculations, we choose 10 for
each, to give a 10$\times$ faster flame speed.  We call this the ``10/10''
flame.  We will also do a simulation with the reactions and
conductivity both boosted by 5, the ``5/5'' flame.

@@

We measure the laminar flame width as:
\begin{equation}
\lambda_L \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}
Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_L/\Delta x \sim 5$~\citep{SNld}.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

@@

To assess the sensitivity of the flame propagation to the various approximations
we made, we ran a suite of simulations. 
Table~\ref{table:sim_names} summarizes these simulations.
The majority of them used a reaction rate boosting of $10$ and a
conductivity boosting of $10$, which should increase the flame speed
by a factor of $10$.  Most simulations used a resolution of
$10~\mbox{cm}$ and a domain width of a little more than one kilometer.  
We use an artifically high rotation rate of
$2000~\mbox{Hz}$, which gives a Rossby length of
\begin{equation}
L_R = \frac{\sqrt{g H_0}}{\Omega} \sim 3\times 10^4~\mathrm{cm}
\end{equation}
using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
quarter of the domain width.  The run at $20~\mbox{cm}$ resolution used
one fewer level of refinement.  The slower rotating case using a
slightly wider domain to accommodate the expected larger Rossby
length.  We note that the entire simulation framework for these
calculations is freely available in the \castro\ github
repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
the simulation names defined in the table to refer to specific runs
and we'll use the 10/10 run as the reference calculation.

\begin{deluxetable}{lcccccc}
\tablecaption{\label{table:sim_names} Simulation parameters.}
\tablehead{\colhead{name} &
           \colhead{reaction} &
           \colhead{conductivity} &
           \colhead{fine grid} &
           \colhead{rotation} &
           \colhead{domain size} &
           \colhead{network} \\
%
           \colhead{} &
           \colhead{boost} &
           \colhead{boost} &
           \colhead{resolution} &
           \colhead{rate} &
           \colhead{} &
           \colhead{}
}
\startdata
10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
10/10-lowres & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
10/10-1000 Hz & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
\enddata
\end{deluxetable}

@@

Our default resolution puts $\sim 80$ zones vertically in the
``cool'' \MarginPar{double check} model atmosphere height.  To
understand the effects of resolution, we also performed a run with one
fewer refinement level, giving a $20~\mbox{cm}$ resolution overall (this is our
10/10-lowres run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
20~ms.  The structure is largely the same as the 10/10 run, with
largely the same flame shape and position, and the same structure in
the energy generation rate.  Since we do not have a jump in refinement
right below the atmosphere, we don't see the temperature feature from
the initial model mapping there, but we do see some cooling in the
\isot{Ni}{56} region.  This is likely a resolution effect, again due
to the strong gradient in temperature at the base of the atmosphere.
The strong agreement in the low resolution run to the 10/10 run gives
us confidence that we are capturing the flame physics properly.

@@

We also considered the effect of the network size.
Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
with the standard 13-isotope {\tt aprox13} network and the reduced
7-isotope {\tt iso7} network.  We see that the flame in the {\tt aprox13}
case is slightly more advanced and has a higher $\bar{A}$ than the {\tt iso7}
case.  We see in the next section that these two networks give largely
the same flame speed.  The reduced network size saves a lot of memory,
which will be useful when we transition to 3D simulations.

\begin{figure}[t]
\centering
\plotone{network_compare}
\caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
  {\tt aprox13} to a run with the {\tt iso7} network.}
\end{figure}

@@

To measure the propagation rate of the burning front, we first collapse our nuclear
energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
profile by averaging over the vertical coordinate. We then take the peak
$\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
We define the position of the flame front to be the location ahead of hottest part of
the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
This corresponds roughly to the leading edge of the burning region. Averaging
over the vertical coordinate helps to reduce sensitivity to localized fluid motions, as does
tracking the $0.1 \%$ contour rather than a local maximum. As we see in Figure
\ref{fig:flame_speed}, the flame settles into a state of steady propagation after an initial
transient period spanning $\sim 3$ ms. The position data here are well fitted by a linear
function of time, and the resulting slope gives the velocity of the flame front.

\begin{deluxetable}{lc}
	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured in 2D calculations.}
	\tablehead{\colhead{run} & \colhead{$s_\mathrm{front}$ (km s$^{-1}$)}} 
	\startdata
	10/10 & $9.18 \pm 0.03$ \\
	5/5 & $4.00 \pm 0.01$ \\
	10/10-iso7 & $7.56 \pm 0.02$ \\
	10/10-lowres & $9.33 \pm 0.04$ \\
	10/10-1000 Hz & $18.6 \pm 0.16$ \\
	\enddata
	%
\end{deluxetable}

Table \ref{table:flame_speeds_multid} gives the flame speed measured in each multidimensional simulation.
The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts (Table \ref{table:flame_speeds_1d}).
The increase in flame speed is likely a product of the larger flame surface area and hydrodynamical effects such as turbulence,
wrinkling, and convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the burning region.

@@

All of the various approximations have the expected effects. The flame in the 5/5
boost run propagates at about half the speed of the 10/10 flame. This is consistent
with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the reduced
Rossby radius \citep{cavecchi:2013}. This confirms that the role of the Coriolis
force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical
effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution had minimal impact on
the flame speed -- the low resolution line is right on top of the fit for the standard run
in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
smaller network also produces similar behavior to the standard run, although there is a small
reduction in speed owing to less energetic burning.

@@

{\color{red} Show whatever it was that Don cooked up.}

Baroclinicity is calculated as
\begin{equation}
    \boldsymbol{\psi} = \frac{1}{\rho^2} \nablab p \times \nablab \rho,
\end{equation}
and shows the misalignment of the local density and pressure gradients
\citep{Malone2014a}. As we are considering a 2D system, the
component of the baroclinicity we consider here (out of the plane)
reflects the misalignment of the fields in the plane of the
simulation. Figure~\ref{fig:baroclinicity} shows that the
baroclinicity peaks along the flame front. This baroclinicity
generates vorticity, which in turn entrains material along the surface
of the flame (see also \citealt{cavecchi:2013}). In 3D, this same vorticity
should perturb the flame front and further increase the flame speed
\citep{Cavecchi2019}.

@@

Vortical motions are also evident in a direct visualization of the velocity field, as seen in
Figure \ref{fig:streamlines}. We observe turbulence in the wake of the flame and in the
unburnt fuel ahead of the front, while a convective cycle sets up near the hottest part of the
burning region. The cycle draws the cooler fluid near the interface into the center of the flame
and drives hot ash out towards the instability at its surface, helping to facilitate the flame
spreading. Convective mixing should still be important in 3D, but we would expect much more
complicated flow patterns, with a greater contribution from small-scale features \citep{xrb3d}.

@@

@@

We use a 2D axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
and Neumann boundaries on all other sides.  In this geometry, the solution behaves like
a spherical problem, and has the form:
\begin{equation}
  T(r,t) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
\end{equation}
where $t_0$ is a small time used to set the initial width of the
Gaussian, $r$ is the distance from the origin, and $D$ is the
diffusion coefficient.  $T_1$ is the ambient temperature and $T_2$ is
the temperature of the peak of the Gaussian.  We use $D = 1$, $T_1 = 1$, $T_2 = 2$,
and $t_0 = 0.001$.
